# INTRODUCTION

salmonMSE is a stochastic decision-support tool for Pacific salmon focusing on strategic trade-offs among harvest, hatchery and habitat management levers.
salmonMSE can be used for risk-based analyses to evaluate the performance of and prioritize management actions and identify trade-offs towards achieving biological and harvest objectives.

salmonMSE is a generalizable tool applicable across salmon species and populations, and expands upon previous simulation tools such as samSim [@holt2020] and AHA [@hsrg2020].
These tools project a simulated population over successive generations to obtain the long-term equilibrium properties of the state dynamics.
To accommodate more complex life histories, salmonMSE is age-structured which is useful for species that have returns comprising multiple brood-years.
Age-structured dynamics also explicitly model juvenile mortality in the marine life stage.
Importantly, salmonMSE was developed to comprehensively support evaluation of harvest policies, hatchery enhancement programs, and changes in early freshwater survival simultaneously.
Finally, stochasticity can be incorporated in the model to evaluate outcomes across uncertain parameters.
In general, its flexibility is intended to encapsulate a gradient of simple to more complex models as data availability to inform parameters allows.

salmonMSE can be used to address questions such as:

- How do mark-selective fisheries impact harvest and conservation objectives?
- How does a change in hatchery release strategy impact harvest and conservation objectives?
- How does a candidate broodtake rule impact the ability to achieve PNI and conservation objectives?
- How do changes in environmental drivers impact ability to rebuild populations?
- What combination of fishery exploitation rates and hatchery production targets meets spawner abundance objectives?

## Structure of document

This document is intended to be summary of the age-structured projection model in salmonMSE.
The model is stochastic, and users have full control of stochastic, time-varying parameters.
For example, users manually sample the desired probability density functions to generate Monte Carlo samples of a parameter and provide them to the model, or specify the year-specific values of a parameter to create a temporal trend in a parameter.
In the interest of simplicity, the equations describe the biological dynamics within a single simulation replicate.

The equations are structured in the following order:

- Juvenile marine abundance in the beginning of the annual time step, followed by preterminal fishing
- Maturation and recruitment of the mature component of the population, followed by terminal fishing
- Escapement from marine fisheries to the in-river return, where there is broodtake for hatchery programs and further in-river removals
- Spawner abundance and egg production
- Survival to outmigrating juveniles of the next generation in the next time step
- Advancement of age classes of juveniles that did not mature in the previous time step

A glossary of important state variables is provided in Table \@ref(tab:glossary). 

Appendices describe various modifications to the system dynamics including mark-selective fishing (Appendix \@ref(app:msf)), fitness impacts on juvenile survival from genetic drift due to selection against hatchery traits in the natural environment (Appendix \@ref(app:fitness)), and multi-population models (Appendix \@ref(app:multi)).

## Source code

The corresponding R code is available on Github at [https://github.com/Blue-Matter/salmonMSE](https://github.com/Blue-Matter/salmonMSE).
`ProjectSOM()` is the primary function that executes the projection and calls subsidiary functions that implement different components of the simulation (Table \@ref(tab:equation-table)).

## Basic inputs and outputs

Users specify the parameters of the simulation through an S4 object of class `SOM`, which in itself is comprised of smaller S4 objects (class Bio, Habitat, Hatchery, or Harvest) that organize the biological dynamics and management levers:

```{r, eval=FALSE, echo=TRUE}
Bio <- new("Bio")
SOM <- new(
  "SOM",
  Bio,
  Habitat,
  Hatchery,
  Harvest
)
```

The simulation of the projection model is run with a call to the `salmonMSE()` function:

```{r, eval=FALSE, echo=TRUE}
SMSE <- salmonMSE(SOM)
```

The output is another S4 object, class SMSE, that reports the results to the projection.
The contents of S4 objects can be assigned or extracted  with the "@" symbol, for example:

```{r, eval=FALSE, echo=TRUE}
Bio@kappa <- 3
SMSE@NOS
```

Table \@ref(tab:glossary) summarizes the important parameters and state variables described in this document, and where they can be assigned from the input or extracted from the output.

\newpage

```{r glossary, results="asis"}
glossary <- readr::read_csv("tables/glossary.csv") %>%
  filter(Type_Internal != "Fitness") %>%
  arrange(Entry) %>%
  select(Name, Definition, Type, Slot)

flextable(glossary) %>%
  compose(j = "Name", value = as_paragraph(as_equation(Name))) %>%
  valign(j = 1:4, valign = "top", part = "all") %>%
  #align(j = 1:4, align = "left", part = "all") %>%
  set_table_properties(layout = "autofit") %>%
  set_caption("Glossary of state variables used in the dynamics equations and their relation to the R code, either as user inputs, internal variables (not reported to user), or reported outputs. Inputs are stored in an S4 object of class SOM and outputs are reported in an S4 object of class SMSE. The table reports the corresponding slots for the state variables in the S4 objects.")
```

\newpage

```{r equation-table, results="asis"} 
#Equations <- read.csv("tables/EquationTable.csv")
#flextable(Equations) |>
#  #ftExtra::colformat_md(j="Eq_Num", metadata = rmarkdown::metadata) |>
#  compose(j = "Eq_Num", value = as_paragraph(Eq_Num)) |>
#  set_table_properties(layout = "autofit")
#can't get references to show up in a table

table_func <- data.frame(
  Description = c("Entire projection", "Juvenile natural production (no habitat modeling)", "Juvenile natural production (with habitat modeling)", "Hatchery production", "Preterminal fishery", "Terminal fishery", "Spawners", "Fitness"),
  Function = c("ProjectSOM()", "nextgen_SRR_func()", "nextgen_habitat_func()", 
               "brood_func()", "catch_func()", "catch_func()", "calc_spawners()", "fitness_func()")
)

flextable(table_func) |>
  set_caption("List of R functions that perform the various calculations.") %>%
  width(j = 1:2, width = 3)
```


