\newpage

# APPENDIX `r new_appendix()`. MIXED BROOD-YEAR PNI {#app:pni}

PNI (proportionate natural influence) is an approximation of the rate of gene flow from the hatchery to the natural environment, calculated for the progeny in year $y+1$ from the parental composition of year $y$:

\begin{align}
 \textrm{PNI}_{y+1} = \dfrac{\sum_a p^{\textrm{NOB}}_{y,a}}{\sum_a p^{\textrm{NOB}}_{y,a} + \sum_a p^{\textrm{HOSeff}}_{y,a}}
 (\#eq:PNI)
\end{align}

Generally, a combination of minimizing hatchery releases, increasing natural broodtake, and reducing the number hatchery origin spawners maintains high PNI, i.e., low rate of gene flow from the hatchery to natural environment.

If there is no natural origin broodtake, e.g., brood is imported or there are only strays, then PNI is calculated with equation 6 of @withler2018:

\begin{align}
 \textrm{PNI}_{y+1} = \dfrac{h^2}{h^2 + (1 - h^2 + \omega^2) \sum_a p^{\textrm{HOSeff}}_{y,a}}
 (\#eq:AllImportPNI)
\end{align}


### Wild salmon

With single brood-year returns, the proportion of wild salmon, natural origin spawners whose parents were also natural spawners, can be calculated as

\begin{align}
  p^\textrm{WILD}_g = (1 - p^\textrm{HOScensus}_g) \times
  \dfrac{(1 - p^\textrm{HOScensus}_{g-1})^2}
  {(1 - p^\textrm{HOScensus}_{g-1})^2 + 2 \gamma \times p^\textrm{HOScensus}_{g-1}(1 - p^\textrm{HOScensus}_{g-1}) +
  \gamma^2 (p^\textrm{HOScensus}_{g-1})^2} \\
  where \\
  p^\textrm{HOScensus} = \textrm{HOS}/(\textrm{HOS} + \textrm{NOS})
  (\#eq:pWild)
\end{align}

The first term is the proportion of natural spawners in the current generation $g$.

The ratio comprising the second term discounts the proportion of the current generation to include natural spawners whose parents were both natural spawners.
Assuming non-assortative mating, the three terms in the denominator gives the composition of generation $g$  whose parents who are both natural origin, mixed origin (one parent in natural origin and the other is hatchery origin), and both hatchery origin.

To generalize for mixed-brood year return, we calculate the probability weighted across brood-years and age class fecundity:

\begin{align}
  p^\textrm{WILD}_y &= \sum_a \dfrac{\textrm{NOS}_{y,a}}{\sum_{a'}(\textrm{NOS}_{y,a'} + \textrm{HOS}_{y,a'})} \times  \\
  & \dfrac{(\sum_{a'} p^\textrm{NOScensus}_{y-a,a'})^2}
  {(\sum_{a'} p^\textrm{NOScensus}_{y-a,a'})^2 + 2 \gamma \times (\sum_{a'}p^\textrm{NOScensus}_{y-a,a'})(\sum_{a'}p^\textrm{HOScensus}_{y-a,a'}) +
  \gamma^2 (\sum_{a'}p^\textrm{HOScensus}_{y-a,a'})^2} \\
  where \\
  p^\textrm{NOScensus}_{y,a} &= \dfrac{\textrm{Fec}_a \times \textrm{NOS}_{y,a}}{\sum_a{\textrm{Fec}_a (\textrm{NOS}_{y,a}} + \textrm{HOS}_{y,a})} \\
  p^\textrm{HOScensus}_{y,a} &= \dfrac{\textrm{Fec}_a \times \textrm{HOS}_{y,a}}{\sum_a{\textrm{Fec}_a (\textrm{NOS}_{y,a}} + \textrm{HOS}_{y,a})}\\
  (\#eq:pWildMixed)
\end{align}


The probability of finding a wild salmon in year $y$ is the sum of probabilities of finding a wild salmon over all ages.
For each age $a$, the first ratio is the probability of finding a natural spawner in year $y$.
The second ratio is the probability of mating success from two parental natural spawners in year $y-a$ using a Punnett square, assuming non-assortative mating across age and origin.
The summation across dummy age variable $a'$ calculates the total proportion of spawners in a given year.

Effective proportions, i.e., weighting by age-class fecundity, in the parental composition accounts for older age classes that are more fecund and more likely to contribute to the production of offspring.

### Mark-selective fishing

If the mark rate $m$ of hatchery fish is greater than zero, then mark-selective fishing can be implemented for either the pre-terminal fishery, terminal fishery, or both with no retention on natural-origin fish.
The mark rate is a proxy for retention and the harvest rate $u^\textrm{harvest}$ corresponds to the kept catch of hatchery-origin fish.
The exploitation rate $u^\textrm{exploit}$ is calculated from kept catch of hatchery-origin fish and dead releases of natural-origin fish.
Exploitation rates differ between hatchery-origin and natural-origin fish because there is no retention of the latter.

Fishing mortality on hatchery-origin fish is partitioned into two parts for kept and release catch:

\begin{align}
  F^\textrm{kept,HO} &= mE\\
  F^\textrm{rel,HO} &= (1 - m)\delta E
  (\#eq:FishMortHatchery)
\end{align}

where $\delta$ is the proportion of released fish that die, i.e., release mortality.

$E$ is an index of fishing effort, also referred to as the encounter rate by the fishery, that links together $F^\textrm{kept,HO}$ and $F^\textrm{rel,NO}$.

Natural-origin fish experience exclusively fishing mortality due to deaths from release:

\begin{align}
  F^\textrm{rel,NO} = \delta E
  (\#eq:FishMortNatural)
\end{align}


Intuitively, fishing effort can increase in a mark-selective fishery compared to a non-selective fishery.
For example, if the mark rate is 20 percent, then the fishing effort could be 500 percent higher than in a non-selective fishery in order to attain the kept quota or bag limit.
Additional catch and release mortality then occurs for un-marked fish, according to $\delta$.

The kept catch $K$ are

\begin{align}
  K^\textrm{HOS,PT}_{y,a} &= \dfrac{F^\textrm{kept,HO,PT}_y}{F^\textrm{kept,HO,PT}_y + F^\textrm{rel,HO,PT}_y}\left(1 - \exp(-v^\textrm{PT}_a[F^\textrm{kept,HO,PT}_y + F^\textrm{rel,HO,PT}_y])\right)N^\textrm{juv,HOS}_{y,a}\\
  K^\textrm{HOS,T}_{y,a} &= \dfrac{F^\textrm{kept,HO,T}_y}{F^\textrm{kept,HO,T}_y + F^\textrm{rel,HO,T}_y}\left(1 - \exp(-v^\textrm{T}_a[F^\textrm{kept,HO,T}_y + F^\textrm{rel,HO,T}_y])\right)\textrm{HOR}_{y,a}
  (\#eq:KeptCatchMS)
\end{align}

If management is by target harvest rate for the preterminal and terminal fisheries, the corresponding $E$ is solved to satisfy:

\begin{align}
  u^\textrm{PT} &= \frac{\sum_aK^\textrm{HOS,PT}_{y,a}\times \textrm{AEQ}_a}{\sum_a(K^\textrm{HOS,PT}_{y,a} \times \textrm{AEQ}_a + N^\textrm{juv,HOS}_{y,a}\exp(-v^\textrm{PT}_a [F^\textrm{kept,HO,PT}_y + F^\textrm{rel,HO,PT}_y])\times r_a)}\\
  u^\textrm{T} &= \frac{\sum_aK^\textrm{HOS,T}_{y,a}}{\textrm{HOR}_{y,a}}
  (\#eq:HarvTargetMS)
\end{align}

If management is by target catch, the corresponding $E$ is solved to satisfy:

\begin{align}
  K^\textrm{PT,target} &= \sum_a K^\textrm{HOS,PT}_{y,a}\\
  K^\textrm{T,target} &= \sum_a K^\textrm{HOS,T}_{y,a}
  (\#eq:CatchTargetMS)
\end{align}

The exploitation rate for natural-origin fish is calculated from dead discards. The exploitation rate for hatchery origin fish is calculated from kept catch and dead discards:

\begin{align}
  u^\textrm{exploit,NOS,PT}_{y,a} &= \frac{\sum_aD^\textrm{NO,PT}_{y,a}\times \textrm{AEQ}_a}{\sum_a(D^\textrm{NO,PT}_{y,a} \times \textrm{AEQ}_a + N^\textrm{juv,NOS}_{y,a}\exp(-v^\textrm{PT}_a F^\textrm{rel,NO,PT}_y)\times r_a)}\\
  u^\textrm{exploit,HOS,PT}_{y,a} &= \frac{\sum_a(D^\textrm{HO,PT}_{y,a} + K^\textrm{HO,PT}_{y,a})\times \textrm{AEQ}_a}{\sum_a((D^\textrm{HO,PT}_{y,a} + K^\textrm{HO,PT}_{y,a}) \times \textrm{AEQ}_a + N^\textrm{juv,HOS}_{y,a}\exp(-v^\textrm{PT}_a [F^\textrm{kept,HO,PT} + F^\textrm{rel,HO,PT}]_y)\times r_a)}\\
  u^\textrm{exploit,NOS,T}_{y,a} &= \frac{\sum_aD^\textrm{NO,T}_{y,a}}{\sum_a\textrm{NOR}_{y,a}}\\
  u^\textrm{exploit,HOS,T}_{y,a} &= \frac{\sum_a(D^\textrm{HO,T}_{y,a} + K^\textrm{HO,T}_{y,a})}{\sum_a\textrm{HOR}_{y,a}}
 (\#eq:ExpRateMS)
\end{align}

where $D$ is dead discards:

\begin{align}
D^\textrm{NO,PT}_{y,a} &= (1 - \exp(-v_a F^\textrm{rel,NO,PT}_y)) N^\textrm{juv,NOS}_{y,a}\\
D^\textrm{HO,PT}_{y,a} &= \frac{F^\textrm{rel,HO,PT}_y}{F^\textrm{kept,HO,PT}_y + F^\textrm{rel,HO,PT}_y}(1 - \exp(-v_a [F^\textrm{kept,HO,PT}_y + F^\textrm{rel,HO,PT}_y]))N^\textrm{juv,HOS}_{y,a}\\
D^\textrm{NO,T}_{y,a} &= (1 - \exp(-v_a F^\textrm{rel,NO,T}_y))\textrm{NOR}_{y,a}\\
D^\textrm{HO,T}_{y,a} &= \frac{F^\textrm{rel,HO,T}_y}{F^\textrm{kept,HO,T}_y + F^\textrm{rel,HO,T}_y}(1 - \exp(-v_a [F^\textrm{kept,HO,T}_y + F^\textrm{rel,HO,T}_y]))\textrm{HOR}_{y,a}
  (\#eq:DiscardsMS)
\end{align}


```{r equation-table, echo = FALSE}
Equations <- read.csv("tables/EquationTable.csv")
```

```{r Equation_Table, tab.cap = "List of equations, and where they exist in the model code."} 
# flextable(Equations) |>
#  # ftExtra::colformat_md(j="Eq_Num", metadata = rmarkdown::metadata) |>
#   compose(j = "Eq_Num", value = as_paragraph(Eq_Num)) |>
#   set_table_properties(layout = "autofit")
#can't get references to show up in a table

table_func <- data.frame(
  Description = c("Entire projection", "Juvenile natural production (no habitat modeling)", "Juvenile natural production (with habitat modeling)", "Hatchery production", "Preterminal fishery", "Terminal fishery", "Spawners", "Fitness"),
  Function = c("ProjectSOM()", "nextgen_SRR_func()", "nextgen_habitat_func()", 
               "brood_func()", "catch_func()", "catch_func()", "calc_spawners()", "fitness_func()")
)

flextable(table_func) |>
  set_caption("List of R functions that perform the various calculations.") %>%
  width(j = 1:2, width = 3)
```


